---
title: "andersfuckoff"
author: "Victor Møller"
date: "13 mar 2019"
output: html_document
---


```{r setup, include=FALSE}

library(pacman)
p_load(rethinking, tidyverse, brms, ggunchained)

# data loading
d <- read.csv("Assignment3Data.csv")

```


#A: 
#preprocessing, scaling

```{r }
#Filtering visit 1 & neurotypicals 
d1 <- d %>%
  filter(Visit == 1) %>%
  filter(ASD == 1)

#centering (JanCo style)
colname <- c("VerbalIQ", "NonVerbalIQ", "SocialIQ")

z_score <- function(var) {
  return( (var - mean(var) ) / sd(var) )
}

d2 <- d1 %>%
  mutate_at(vars(colname), z_score)

#standardizing 
d2$ADOS <- d2$ADOS / sd(d2$ADOS)

```


#B

BRMS --> DON't RUN THIS CHUNK IT TAKES AGES. 
```{r}
# making a formula, automatic prior detection
m1 <- bf(ADOS ~ 1 + NonVerbalIQ)
get_prior(m1, d2, family = gaussian) #outrageusly high it seems?

m2 <- bf(ADOS ~ 1 + VerbalIQ)
get_prior(m2, d2, family = gaussian) #outrageously high. 

m3 <- bf(ADOS ~ 1 + SocialIQ)
get_prior(m3, d2, family = gaussian) #outrageously high. 


#prior predictive check - Riccardo style 
prior = c(prior(normal(3, 1), class = "Intercept"), 
          prior(normal(-0.5, 0.5), class = "b"), 
          prior(normal(1,0.5), class = "sigma"))


#model with = "only" (commented bc. of run-time)
m1_only <- brm(m1, family = gaussian, d2, 
                prior = prior, sample_prior = "only", 
                chain = 1, iter = 500)  

m2_only <- brm(m2, family = gaussian, d2, 
                prior = prior, sample_prior = "only", 
                chain = 1, iter = 500)

m3_only <- brm(m3, family = gaussian, d2, 
                prior = prior, sample_prior = "only", 
                chain = 1, iter = 500)

#plots prior? 
plot(m1_only)
plot(m2_only)
plot(m3_only)

#model with = TRUE
m1_true <- brm(m1, family = gaussian, d2, 
               prior = prior, sample_prior = TRUE, #posterior?
               chain = 1, iter = 500)

m2_true <- brm(m2, family = gaussian, d2, 
               prior = prior, sample_prior = TRUE, 
               chain = 1, iter = 500)

m3_true <- brm(m3, family = gaussian, d2, 
               prior = prior, sample_prior = TRUE, 
               chain = 1, iter = 500)

#plots posterior?
plot(m1_true)
plot(m2_true)
plot(m3_true)

#pp_check
pp_check(m1_only) #looks ok. 
pp_check(m2_only) #looks ok. 
pp_check(m3_only) #looks ok. 
pp_check(m1_true) #looks good. 
pp_check(m2_true) #somewhat shifted?
pp_check(m3_true) #looks good. 

```

#C:
#Doing the 3 models 

```{r}
library(rethinking)

#making model (getting help from glimmer)
glimmer(ADOS ~ 1 + NonVerbalIQ, d2)

#model 1
nonverbal <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*NonVerbalIQ,
    a ~ dnorm(3, 1), 
    b ~ dnorm(-0.5, 0.5),
    sigma ~ dnorm(1, 0.5)
    ),
  data = d2)

precis(nonverbal, corr = TRUE) 

#model 2
verbal <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*VerbalIQ,
    a ~ dnorm(3, 1), 
    b ~ dnorm(-0.5, 0.5),
    sigma ~ dnorm(1, 0.5)
    ),
  data = d2)

precis(verbal, corr = TRUE) 

#model 3
social <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*SocialIQ,
    a ~ dnorm(3, 1), 
    b ~ dnorm(-0.5, 0.5),
    sigma ~ dnorm(1, 0.5)
    ),
  data = d2)

precis(social, corr = TRUE) 


```

results of models. 
I.e., showing precis outputs. 

```{r}

plot(precis(nonverbal))
plot(precis(verbal))
plot(precis(social))

```

assessing model quality for nonverbal 

```{r}

summary(d2$NonVerbalIQ)
summary(d2$ADOS)

NonVerbal.seq <- seq(from = -3, to = 3, by = 1)
sim.NonVerbal <- sim(nonverbal, data = list(NonVerbalIQ = NonVerbal.seq))
ADOS.PI <- apply(sim.NonVerbal, 2, PI, prob = 0.89)
mu <- link(nonverbal)
mu <- link(nonverbal, data = data.frame(NonVerbalIQ = NonVerbal.seq))
mu.mean <- apply(mu, 2, mean) 
mu.HPDI <- apply(mu, 2, HPDI, prob = 0.89)

plot(ADOS ~ NonVerbalIQ, d2, col = col.alpha(rangi2, 1))
lines(NonVerbal.seq, mu.mean)
# draw HPDI region for line
shade(mu.HPDI, NonVerbal.seq )
# draw PI region for simulated heights
shade(ADOS.PI, NonVerbal.seq ) 

```

assesing model for verbal

```{r}
summary(d2$VerbalIQ)
summary(d2$NonVerbalIQ)

Verbal.seq <- seq(from = -2, to = 2, by = 1)
sim.Verbal <- sim(verbal, data = list(VerbalIQ = Verbal.seq))
ADOS.PI <- apply(sim.Verbal, 2, PI, prob = 0.89)
mu <- link(verbal, data = data.frame(VerbalIQ = Verbal.seq))
mu.mean <- apply(mu, 2, mean) 
mu.HPDI <- apply(mu, 2, HPDI, prob = 0.89)

plot(ADOS ~ VerbalIQ, d2, col = col.alpha(rangi2, 1))
lines(Verbal.seq, mu.mean)
# draw HPDI region for line
shade(mu.HPDI, Verbal.seq )
# draw PI region for simulated heights
shade(ADOS.PI, Verbal.seq ) 
```

assessing model for social

```{r}
summary(d2$SocialIQ)

Social.seq <- seq(from = -2, to = 3, by = 1)
sim.Social <- sim(social, data = list(SocialIQ = Social.seq))
ADOS.PI <- apply(sim.Social, 2, PI, prob = 0.89)
mu <- link(social, data = data.frame(SocialIQ = Social.seq))
mu.mean <- apply(mu, 2, mean) 
mu.HPDI <- apply(mu, 2, HPDI, prob = 0.89)

plot(ADOS ~ SocialIQ, d2, col = col.alpha(rangi2, 1))
lines(Social.seq, mu.mean)
# draw HPDI region for line
shade(mu.HPDI, Social.seq )
# draw PI region for simulated heights
shade(ADOS.PI, Social.seq ) 

```

TASK 2: 
#A different portions of variance?

```{r}
#corr-check 
d2cor <- d2[6:8] #weird ones 
cor(d2cor) #social & nonverbal are not that correlated. 

```

#B all IQ measures in same model? 

```{r}
mNonV <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_NonVIQ*NonVerbalIQ,
Intercept ~ dnorm(3, 1),
b_NonVIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = d2)

mV <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_VIQ*VerbalIQ,
Intercept ~ dnorm(3, 1),
b_VIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = d2)

MS <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_SIQ*SocialIQ,
Intercept ~ dnorm(3, 1),
b_SIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = d2)

m1 <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_NonVIQ*NonVerbalIQ + b_VIQ * VerbalIQ + b_SIQ * SocialIQ,
Intercept ~ dnorm(3, 1),
b_NonVIQ ~ dnorm(-0.5, 0.5),
b_VIQ ~ dnorm(-0.5, 0.5), 
b_SIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = d2)

m2 <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_NonVIQ*NonVerbalIQ + b_SIQ * SocialIQ,
Intercept ~ dnorm(3, 1),
b_NonVIQ ~ dnorm(-0.5, 0.5),
b_SIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = d2)

precis(mNonV)
precis(mV)
precis(MS)
precis(m1, cor = TRUE)
precis(m2, cor = TRUE)

set.seed(1337)
model_comp <- compare(mNonV, mV, MS, m1, m2, n = 10^4)
model_comp #best = 

```

2C # Assessing quality of predictors. 
First block is for non-verbal.

```{r}
# mean & spread. 
summary(d2$SocialIQ)
summary(d2$NonVerbalIQ) #0
summary(d2$VerbalIQ) #0

non.seq <- seq(from=-2.5,to=3,length.out=30)
d.predict <- list(
ADOS = rep(0,30), # empty outcome 
NonVerbalIQ = non.seq, # sequence of social IQ
SocialIQ = rep(0,30), # average of non-verbal
VerbalIQ = rep(0,30) #average of verbal 
)

pred.m1 <- link( m1 , data=d.predict )
mu <- apply( pred.m1 , 2 , mean )
mu.PI <- apply( pred.m1 , 2 , PI )
# plot it all
plot( ADOS ~ NonVerbalIQ , d2 , col=rangi2 )
lines( non.seq , mu , lty=2 )
lines( non.seq , mu.PI[1,] , lty=2 )
lines( non.seq , mu.PI[2,] , lty=2 )

#ensemble
IQ.ensemble <- ensemble( mNonV, m2, m1, data=d.predict ) #also models without?
IQ.ensemble <- ensemble(mNonV, mV, MS, m1, m2, data = d.predict)
mu <- apply( IQ.ensemble$link , 2 , mean )
mu.PI <- apply( IQ.ensemble$link , 2 , PI )
lines( non.seq , mu )
shade( mu.PI , non.seq )

```

2C # Assessing quality of predictors. 
This block is for verbal 

```{r}
# mean & spread. 
summary(d2$SocialIQ)
summary(d2$NonVerbalIQ) #0
summary(d2$VerbalIQ) #0

verb.seq <- seq(from=-1.5,to=2,length.out=30)
d.predict <- list(
ADOS = rep(0,30), # empty outcome 
VerbalIQ = verb.seq, # sequence of social IQ
SocialIQ = rep(0,30), # average of non-verbal
NonVerbalIQ = rep(0,30) #average of verbal 
)

pred.m1 <- link( m1 , data=d.predict )
mu <- apply( pred.m1 , 2 , mean )
mu.PI <- apply( pred.m1 , 2 , PI )
# plot it all
plot( ADOS ~ VerbalIQ , d2 , col=rangi2 )
lines( verb.seq , mu , lty=2 )
lines( verb.seq , mu.PI[1,] , lty=2 )
lines( verb.seq , mu.PI[2,] , lty=2 )

#ensemble
IQ.ensemble <- ensemble(mNonV, mV, MS, m1, m2, data = d.predict)
mu <- apply( IQ.ensemble$link , 2 , mean )
mu.PI <- apply( IQ.ensemble$link , 2 , PI )
lines( verb.seq , mu )
shade( mu.PI , verb.seq )

```

2C # Assessing quality of predictors. 
This block is for social

```{r}
# mean & spread. 
summary(d2$SocialIQ)
summary(d2$NonVerbalIQ) #0
summary(d2$VerbalIQ) #0

soc.seq <- seq(from=-1.5,to=2.5,length.out=30)
d.predict <- list(
ADOS = rep(0,30), # empty outcome 
SocialIQ = soc.seq, # sequence of social IQ
VerbalIQ = rep(0,30), # average of non-verbal
NonVerbalIQ = rep(0,30) #average of verbal 
)

pred.m1 <- link( m1 , data=d.predict )
mu <- apply( pred.m1 , 2 , mean )
mu.PI <- apply( pred.m1 , 2 , PI )
# plot it all
plot( ADOS ~ SocialIQ , d2 , col=rangi2 )
lines( soc.seq , mu , lty=2 )
lines( soc.seq , mu.PI[1,] , lty=2 )
lines( soc.seq , mu.PI[2,] , lty=2 )

#ensemble
IQ.ensemble <- ensemble(mNonV, mV, MS, m1, m2, data = d.predict)
mu <- apply( IQ.ensemble$link , 2 , mean )
mu.PI <- apply( IQ.ensemble$link , 2 , PI )
lines( soc.seq , mu )
shade( mu.PI , soc.seq )
```

TASK 3 - making another model just for TD.
Although, I am not sure how this helps us answer the question

```{r}
#without interaction
d3 <- d %>%
  filter(Visit == 1 & ASD == 0)
colname <- c("VerbalIQ", "NonVerbalIQ", "SocialIQ")
d4 <- d3[complete.cases(d3[6:8]) , ]
d4 <- d4 %>%
  mutate_at(vars(colname), z_score)
d4$ADOS <- d4$ADOS / sd(d4$ADOS)

#model 1
mean(d4$ADOS) # find 

m1TD <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_NonVIQ*NonVerbalIQ + b_VIQ * VerbalIQ + b_SIQ * SocialIQ,
Intercept ~ dnorm(0.5, 0.5),
b_NonVIQ ~ dnorm(-0.5, 1),
b_VIQ ~ dnorm(-0.5, 1), 
b_SIQ ~ dnorm(-0.5, 1),
sigma ~ dnorm(1, 0.5)
),
data = d4)

precis(m1TD)

```

TASK 3:
visually inspecting whether IQ scores make sense for 
ASD and TD children respectively. 
Specifically checking for interaction effect (which is there..)

```{r}

#Is there an interaction 
d5 <- rbind(d2, d4)
d5$ASD <- as.factor(d5$ASD)

#checking with plots 
library(tidyverse)
p1 <- ggplot(d5, aes(NonVerbalIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by non-verbal IQ for ASD and TD")

p2 <- ggplot(d5, aes(VerbalIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by verbal IQ for ASD and TD")

p3 <- ggplot(d5, aes(SocialIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by social IQ for ASD and TD")

library(cowplot)
plot_grid(p1, p2, p3)

```

Model with interaction
https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/interactions.html

```{r}
#getting an interaction in alist. 
glimmer(ADOS ~ 1 + NonVerbalIQ * ASD, d5)

#stealing from Knaehoej Karse. 
interaction <- bf(ADOS ~ 1 + ASD * SocialIQ)
get_prior(interaction, d5)

#running: notice that I have just copied priors so they suck. 
prior <- c(prior(normal(0,2.5), class = "Intercept"), prior(normal(-0.4, 0.2), class = "b"), prior(normal(0, 7), class = "sigma"))

#model priors. 
model_prior <- brm(interaction, family = gaussian, d5, prior=prior, sample_prior = "only")

#checking. 
summary(d5$ADOS) #not mean centered (knaehoej's is, so again, priors tweaked)
pp_check(model_prior, nsamples = 100)
plot(model_prior)

#this is going to be bullshit because priors are bad, but the code makes sense (maybe)
model1 <- brm(interaction, family = gaussian, d5, prior=prior, sample_prior = T)
pp_check(model1)
summary(model1) #gets an interaction (actually significant?)
```



TASK 3 (old?)

```{r}

#Filtering visit 1 & neurotypicals 
d3 <- d %>%
  filter(Visit == 1) 

#Centering (JanCo style, using z-score)
colname <- c("VerbalIQ", "NonVerbalIQ", "SocialIQ")
d4 <- d3[complete.cases(d3[6:8]) , ]
d4 <- d4 %>%
  mutate_at(vars(colname), z_score)
d4$ADOS <- d4$ADOS / sd(d4$ADOS)

#corr-check 
d2cor <- d2[6:8] #weird ones 
cor(d2cor)

d4cor <- d4[6:8] #both. 
cor(d4cor)

#
lm(ADOS ~ SocialIQ, data = d2)
lm(ADOS ~ NonVerbalIQ, data = d2)
lm(ADOS ~ VerbalIQ, data = d2)

lm(ADOS ~ SocialIQ, data = d4)
lm(ADOS ~ NonVerbalIQ, data = d4)
lm(ADOS ~ VerbalIQ, data = d4)
```



```{r}
dcomb <- rbind(d2, d4)

mNonVTD <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_NonVIQ*NonVerbalIQ,
Intercept ~ dnorm(3, 1),
b_NonVIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = dcomb)

mVTD <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_VIQ*VerbalIQ,
Intercept ~ dnorm(3, 1),
b_VIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = dcomb)

MSTD <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_SIQ*SocialIQ,
Intercept ~ dnorm(3, 1),
b_SIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = dcomb)

m1TD <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_NonVIQ*NonVerbalIQ + b_VIQ * VerbalIQ + b_SIQ * SocialIQ,
Intercept ~ dnorm(3, 1),
b_NonVIQ ~ dnorm(-0.5, 0.5),
b_VIQ ~ dnorm(-0.5, 0.5), 
b_SIQ ~ dnorm(-0.5, 0.5),
sigma ~ dnorm(1, 0.5)
),
data = dcomb)

precis(mNonVTD)
precis(mVTD)
precis(MSTD)
precis(m1TD)
```

TASK 5

```{r}

#model 3
social <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*SocialIQ,
    a ~ dnorm(3, 1), 
    b ~ dnorm(-0.5, 0.5),
    sigma ~ dnorm(1, 0.5),
    ADOS_error ~ dnorm(ADOS, ADOS_SD),
    ADOS_SD <- 0.5
    ),
  data = d2)



```

